Verification of a Deterministic Random Bits Generator

### 2015-12-31 paper/paper.tex
updated instantiate function in paper

The \lstinline{instantiate} function initializes the DRBG state with randomness from a source of entropy. The caller is responsible for specifying the minimum security strength for this DRBG, which for an HMAC-based DRBG translates to “how many bits of entropy should be requested?”. In addition to entropy, the DRBG also asks for a nonce, which is roughly defined as a string of bits with at least \lstinline{1/2 security_strength} bits of entropy. In fact, this nonce can even be taken from the same source of entropy. For the purposes of this project, I am choosing to specialize the specification so that the nonce is \textit{always} taken from the same source of entropy. This is compatible with \stdtitle{mbedTLS}.

@@

\begin{lstlisting}
    Definition DRBG_instantiate_function
      (instantiate_algorithm: list Z -> list Z -> list Z -> Z -> DRBG_working_state)
      (min_entropy_length max_entropy_length: Z)
      (get_nonce: unit -> list Z)
      (highest_supported_security_strength: Z)
      (max_personalization_string_length: Z)
      (prediction_resistance_supported: bool)
      (entropy_stream: ENTROPY.stream)
      (requested_instantiation_security_strength: Z)
      (prediction_resistance_flag: bool)
      (personalization_string: list Z)
      : ENTROPY.result DRBG_state_handle :=
      if requested_instantiation_security_strength >? highest_supported_security_strength
        then ENTROPY.error ENTROPY.generic_error entropy_stream
      else match prediction_resistance_flag, prediction_resistance_supported with
             | true, false => ENTROPY.error ENTROPY.generic_error entropy_stream
             | _,_ =>
               if (Zlength personalization_string) >? max_personalization_string_length
                 then ENTROPY.error ENTROPY.generic_error entropy_stream
               else
                 let security_strength :=
                   if requested_instantiation_security_strength <=? 14 then Some 14
                   else if requested_instantiation_security_strength <=? 16 then Some 16
                   else if requested_instantiation_security_strength <=? 24 then Some 24
                   else if requested_instantiation_security_strength <=? 32 then Some 32
                   else None in
                 match security_strength with
                   | None => ENTROPY.error ENTROPY.generic_error entropy_stream
                   | Some security_strength =>
                   match get_entropy security_strength min_entropy_length max_entropy_length prediction_resistance_flag entropy_stream with
                     | ENTROPY.error e s' => ENTROPY.error ENTROPY.catastrophic_error s'
                     | ENTROPY.success entropy_input entropy_stream =>
                       match get_entropy (security_strength/2) (min_entropy_length/2) (max_entropy_length/2) prediction_resistance_flag entropy_stream with
                         | ENTROPY.error e s' => ENTROPY.error ENTROPY.catastrophic_error s'
                         | ENTROPY.success nonce entropy_stream =>
                           let initial_working_state := instantiate_algorithm entropy_input nonce
                             personalization_string security_strength in
                           ENTROPY.success (initial_working_state, security_strength,
                             prediction_resistance_flag) entropy_stream
                       end
                   end
                 end
           end.
\end{lstlisting}

>>>>

The \lstinline{instantiate} function initializes the DRBG state with randomness from a source of entropy. The caller is responsible for specifying the minimum security strength for this DRBG, which for an HMAC-based DRBG translates to “how many bits of entropy should be requested?”. In addition to entropy, the DRBG also asks for a nonce, which is roughly defined as a string of bits with at least \lstinline{1/2 security_strength} bits of entropy. In fact, this nonce can even be taken from the same source of entropy. In the functional specification, I am supporting both behaviors by having a parameter \lstinline{provided_nonce: option (list Z)} that is used as the nonce if a value is given. In the case that \lstinline{provided_nonce} is \lstinline{None}, the nonce is taken from the entropy source.

@@

\begin{lstlisting}
    Definition DRBG_instantiate_function
      (instantiate_algorithm: list Z -> list Z -> list Z -> Z -> DRBG_working_state)
      (min_entropy_length max_entropy_length: Z)
      (provided_nonce: option (list Z))
      (highest_supported_security_strength: Z)
      (max_personalization_string_length: Z)
      (prediction_resistance_supported: bool)
      (entropy_stream: ENTROPY.stream)
      (requested_instantiation_security_strength: Z)
      (prediction_resistance_flag: bool)
      (personalization_string: list Z)
      : ENTROPY.result DRBG_state_handle :=
        if requested_instantiation_security_strength >? highest_supported_security_strength
          then ENTROPY.error ENTROPY.generic_error entropy_stream
        else match prediction_resistance_flag, prediction_resistance_supported with
               | true, false => ENTROPY.error ENTROPY.generic_error entropy_stream
               | _,_ =>
                 if (Zlength personalization_string) >? max_personalization_string_length then ENTROPY.error ENTROPY.generic_error entropy_stream
                 else
                   let security_strength :=
                     if requested_instantiation_security_strength <=? 14 then Some 14
                     else if requested_instantiation_security_strength <=? 16 then Some 16
                     else if requested_instantiation_security_strength <=? 24 then Some 24
                     else if requested_instantiation_security_strength <=? 32 then Some 32
                     else None in
                   match security_strength with
                     | None => ENTROPY.error ENTROPY.generic_error entropy_stream
                     | Some security_strength =>
                     match get_entropy security_strength min_entropy_length max_entropy_length prediction_resistance_flag entropy_stream with
                       | ENTROPY.error e s' => ENTROPY.error ENTROPY.catastrophic_error s'
                       | ENTROPY.success entropy_input entropy_stream =>
                         let nonce_result :=
                           match provided_nonce with
                             | Some n => ENTROPY.success n entropy_stream
                             | None => get_entropy (security_strength/2) (min_entropy_length/2) (max_entropy_length/2) prediction_resistance_flag entropy_stream
                           end in
                         match nonce_result with
                           | ENTROPY.error e s' => ENTROPY.error ENTROPY.catastrophic_error s'
                           | ENTROPY.success nonce entropy_stream =>
                             let initial_working_state := instantiate_algorithm entropy_input nonce personalization_string security_strength in
                             ENTROPY.success (initial_working_state, security_strength, prediction_resistance_flag) entropy_stream
                         end
                     end
                   end
             end.
\end{lstlisting}
